Measurement Clusterization in D-optimal Designs for Bayesian Linear Inverse Problems over Hilbert Spaces

### 2023-05-22 arxiv/ms.tex
With supp material in arxiv lib

\address{Porter School of the Environment and Earth Sciences, Tel Aviv
  University\\ Tel Aviv, Israel}
\address{Sackler Faculty of Medicine,
  Sciences, Tel Aviv University\\ Tel Aviv, Israel}

@@

\begin{abstract}
  Inversion for parameters of a physical process typically requires
  taking expensive measurements, and the task of finding an optimal
  set of measurements is known as the optimal design
  problem. Surprisingly, measurement locations in optimal designs are
  sometimes extremely clustered, and since measurement clusterization
  is nonintuitive, researchers often avoid it by modifying the optimal
  design problem. We consider a certain flavor of the optimal design
  problem, based on the Bayesian D-optimality criterion, and suggest
  an analytically tractable model for D-optimal designs in Bayesian
  linear inverse problems over Hilbert spaces. We demonstrate that
  measurement clusterization is a generic property of D-optimal
  designs, and prove that correlated noise between measurements
  mitigates clusterization. We also give a full characterization of
  D-optimal designs, which shows that D-optimal designs uniformly
  reduce uncertainty in a select subset of prior covariance
  eigenvectors. Finally, we show how measurement clusterization is a
  consequence of the characterization mentioned above and the
  pigeonhole principle.
\end{abstract}
\maketitle

>>>>

\address{Porter School of the Environment and Earth Sciences, Tel Aviv
  University\\ Tel Aviv, Israel}
%
\address{Sackler Faculty of Medicine, Tel Aviv University\\ Tel Aviv,
  Israel}

@@

\begin{abstract}
  Inversion for parameters of a physical process typically requires taking expensive measurements, and the task of finding an optimal set of measurements is known as the optimal design problem. Surprisingly, measurement locations in optimal designs are sometimes extremely clustered, and researchers often avoid measurement clusterization by modifying the optimal design problem. We consider a certain flavor of the optimal design problem, based on the Bayesian D-optimality criterion, and suggest an analytically tractable model for D-optimal designs in Bayesian linear inverse problems over Hilbert spaces. We demonstrate that measurement clusterization is a generic property of D-optimal designs, and prove that correlated noise between measurements mitigates clusterization. We also give a full characterization of D-optimal designs under our model: We prove that D-optimal designs uniformly reduce uncertainty in a select subset of prior covariance eigenvectors. Finally, we show how measurement clusterization is a consequence of the characterization mentioned above and the pigeonhole principle.
\end{abstract}
\maketitle
